---
title: "The Rise and Fall of 'Shania'"
subtitle: "A 100-Year Pop Culture Case Study in Baby Names"
author: "Shania"
date: today
format:
  html:
    theme: pulse
    toc: true
    smooth-scroll: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
```

```{r}
#| label: load-data
# List all yearly baby name text files in the data directory
files <- list.files("data", pattern = "^yob.*\\.txt$", full.names = TRUE)

# Read all files into a single tibble and extract the year from the filename
baby_names <- read_csv(
  files,
  id = "file",
  col_names = c("name", "sex", "n"),
  show_col_types = FALSE
) |>
  mutate(year = parse_number(basename(file)))
```

## Recent Trend (Last 10 Years)

Over the last 10 years, the popularity of the name **Shania** has generally trended downward, decreasing from 128 babies in 2016 to 70 babies in 2025.

```{r}
#| label: plot-shania
baby_names |>
  filter(name == "Shania", year >= max(year) - 10) |>
  group_by(year) |>
  summarize(total = sum(n)) |>
  ggplot(aes(x = year, y = total)) +
  geom_area(fill = "#c084fc", alpha = 0.25) +
  geom_line(color = "#7e22ce", linewidth = 1.2) +
  geom_point(color = "#581c87", size = 2.5) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Popularity of the Name 'Shania' (Last 10 Years)",
    subtitle = "A steady decline as 1990s pop-culture influence fades into history",
    caption = "Source: US Social Security Administration",
    x = "Year",
    y = "Number of Babies"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14, color = "#581c87"),
    plot.subtitle = element_text(color = "gray30", margin = margin(b = 10))
  )
```

## 100-Year Historical View

Over the last 100 years, the name **Shania** was virtually nonexistent in records until 1970. Its dramatic explosion in the mid-1990s directly aligns with the February 1995 release of Shania Twain's smash-hit album, *[The Woman in Me](https://en.wikipedia.org/wiki/The_Woman_in_Me_(album))*. Following this all-time peak of 1,837 babies in 1996, the name has steadily returned toward baseline.

```{r}
#| label: plot-shania-100
baby_names |>
  filter(name == "Shania", year >= max(year) - 100) |>
  group_by(year) |>
  summarize(total = sum(n)) |>
  ggplot(aes(x = year, y = total)) +
  geom_area(fill = "#c084fc", alpha = 0.25) +
  geom_line(color = "#7e22ce", linewidth = 1.2) +
  geom_vline(xintercept = 1995, linetype = "dashed", color = "#7e22ce", linewidth = 0.8) +
  annotate(
    "text",
    x = 1995,
    y = 1750,
    label = "1995: 'The Woman in Me' released",
    hjust = 1.05,
    size = 3.8,
    fontface = "italic",
    color = "#581c87"
  ) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Popularity of the Name 'Shania' Over the Last 100 Years",
    subtitle = "Surged following the 1995 release of Shania Twain's breakthrough album",
    caption = "Source: US Social Security Administration",
    x = "Year",
    y = "Number of Babies"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14, color = "#581c87"),
    plot.subtitle = element_text(color = "gray30", margin = margin(b = 10))
  )
```
