---
title: "Baby Names"
format: html
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
library(tidyverse)
```

```{r}
#| label: load-data
# List all yearly baby name text files in the data directory
files <- list.files("data", pattern = "^yob.*\\.txt$", full.names = TRUE)

# Read all files into a single tibble and extract the year from the filename
baby_names <- read_csv(
  files,
  id = "file",
  col_names = c("name", "sex", "n"),
  show_col_types = FALSE
) |>
  mutate(year = parse_number(basename(file)))
```

Over the last 10 years, the popularity of the name **Shania** has generally trended downward, decreasing from 128 babies in 2016 to 70 babies in 2025.

```{r}
#| label: plot-shania
baby_names |>
  filter(name == "Shania", year >= max(year) - 10) |>
  group_by(year) |>
  summarize(total = sum(n)) |>
  ggplot(aes(x = year, y = total)) +
  geom_line(color = "steelblue", linewidth = 1) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Popularity of the Name 'Shania' (Last 10 Years)",
    x = "Year",
    y = "Number of Babies"
  ) +
  theme_minimal()
```

Over the last 100 years, the name **Shania** was absent from records until 1970. It experienced a dramatic surge in the mid-to-late 1990s, peaking in 1996 with 1,837 babies (likely influenced by country star Shania Twain), followed by a steady decline over the last two decades.

```{r}
#| label: plot-shania-100
baby_names |>
  filter(name == "Shania", year >= max(year) - 100) |>
  group_by(year) |>
  summarize(total = sum(n)) |>
  ggplot(aes(x = year, y = total)) +
  geom_line(color = "darkred", linewidth = 1) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Popularity of the Name 'Shania' Over the Last 100 Years",
    x = "Year",
    y = "Number of Babies"
  ) +
  theme_minimal()
```
